---
title: "Optimizing Precision and Power by Covariate Selection in Randomized Trials with an Application to COVID-19"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Nicholas Williams, Michael Rosenblum, Iván Díaz

------------------------------------------------------------------------

This repository contains the R code for the simulation studies in the paper "Optimizing Precision and Power by Covariate Selection in Randomized Trials with an Application to COVID-19."

## Directory guide

The `R/` directory contains the source code for functions used in the simulations. The `scripts/` directory contains R and bash scripts for cleaning the raw data, calculating true values, running the simulations on a computing cluster, and turning simulation results into LaTeX tables. 

## Simulation data

The COVID-19 data used in the paper will not be shared. However, synthetic versions of the data have been made available in the `data/public` directory.  

## adjrct

The estimators described in the paper are implemented with a special version of the *adjrct* package. The source code for the package is included with this repository as a .tar.gz file. The package can be installed with `install.packages("adjrct_0.1.0.9000.tar.gz", type = "source", repos = NULL)`.

## Running simulations

Simulations are intended to be run on a high performance computing cluster using Slurm. `scripts/simulate.sh` establishes the computing parameters (this file contains code specific to the authors machines and will need to be modified accordingly by other users) and runs `scripts/simulate.R`. `scripts/simulate.R` uses configuration parameters defined in `scripts/config.yml` to execute different simulation scenarios.

## Tutorial

#### Loading data

```{r}
box::use(dgm = ./R/data)

synthetic <- dgm$covid("ordinal", private = FALSE)
```

### Generating RCT data

```{r}
tmp <- dgm$generate_data(synthetic, "ordinal", prognostic = TRUE, seed = 123, n = 500, effect_size = 3)
head(tmp)
```

### Running a simulation

The `simulate()` function runs a single iteration of a scenario. In the following example, we run an instance for an ordinal outcome in an RCT with 500 observations where we adjust for age, sex, dyspnea, and supplemental o2 use. These covariates are prognostic of the outcome; variable selection is performed using $\ell1$ regularization. 

```{r}
box::use(sim = ./R/simulate)

sim$simulate(synthetic, type = "ordinal", covar = c("age", "sex", "dyspnea", "o2"), 
             prognostic = TRUE, seed = 123, n = 500, effect_size = 3, algo = "lasso", 
             crossfit = FALSE)
```

The `partition()` function takes a data frame of simulation parameters where each row corresponds to a single iteration and dispatches these scenarios to the `simulate()` function.
